VM-Centric Snapshot Deduplication for Converged Cloud Architectures

### 2014-07-19 snapshot_system/paper_del.tex
re

Snapshot deletions can occur frequently since old snapshots become less useful.
Both space accounting and reclamation are difficult because deduplication
greatly complicates how we identify unused data when deletion occurs.
To filter out un-referenced data from shared duplicates,
it would require to either maintain expensive live schemas or conduct
global reference counting~\cite{Guo2011}.
%or use some advanced mark-and-sweep methodologies
%which are all unsuitable for converged architecture as discussed in section~\ref{sect:background}.
% to identify if a chunk can be safely removed without any reference.
%Since our VM-centric design restricts sharing of data chunks only to a small dataset,
%we can greatly simplify the deletion process by 
%focusing on  unreferenced non-PDS chunks within each VM. 
%This process can be conducted independently per VM, and thus results in  a simpler flow control and 
%lower resource usage. 
One of the key benefits of our VM-centric design is that it allows for faster and simpler deletion. 
We propose the following strategies to speedup the deletion process and reduce its resource usage.
\begin{itemize}
\item First, since
sharing in our approach is limited to only a small dataset, we can greatly simplify deletion by focusing
on intra-VM sharing rather than inter-VM sharing. Then the snapshot
 deletion can therefore be conducted independently among VMs.
\item Second,  we propose an {\em approximate} deletion strategy to trade deletion accuracy for speed and
resource usage. Since the VM size is highly skewed in practice,  a large VM may still require
a substantial amount of memory  to facilitate the mark-sweep process of data chunks used within all snapshots
of this VM. 
Our approximate deletion method sacrifices a small percent of storage leakage
to efficiently identify unused chunks using a per-snapshot bloom-filter.
\end{itemize}
The simplified deletion algorithm contains three aspects.
\begin{itemize}
\item {\bf Computation for snapshot reference summary.}
Every time there is a new snapshot created,
we compute a Bloom-filter with $z$ bits as the reference summary vector for all non-PDS chunks used 
in this snapshot.
The items we put into the summary vector are all the references appearing in the metadata of the snapshot.
For each VM we preset the vector size according to estimated VM image size;
given $h$ snapshots stored for a VM, there are $h$ summary vectors maintained.
We adjust the summary vector size and recompute the vectors if the VM size changes substantially over time.
This can be done during periodic leakage repair described below.

>>>>

Snapshot deletions can occur frequently since old snapshots become less useful.
Both space accounting and reclamation are difficult because deduplication
greatly complicates the  identification of unused data when deletion occurs.
To filter out un-referenced data from shared duplicates,
it would require to either maintain expensive live schemas or conduct
global reference counting~\cite{Guo2011}.
%or use some advanced mark-and-sweep methodologies
%which are all unsuitable for converged architecture as discussed in section~\ref{sect:background}.
% to identify if a chunk can be safely removed without any reference.
%Since our VM-centric design restricts sharing of data chunks only to a small dataset,
%we can greatly simplify the deletion process by 
%focusing on  unreferenced non-PDS chunks within each VM. 
%This process can be conducted independently per VM, and thus results in  a simpler flow control and 
%lower resource usage. 
For example, the experiment in Section~\ref{sect:evaldel} shows that
 the grouped mark-and-sweep process takes 43 hours for a 2500 VM cluster using upto 3GB memory per machine.

One of the key benefits of our VM-centric design is that it allows for faster and simpler deletion. 
We propose the following strategies to speedup the deletion process and reduce its resource usage.
\begin{itemize}
\item First, reference counting is focused on non-PDS chunks which are not shared among VMs,
and the mark-and-sweep process is applied within each VM. 
Therefore,  the snapshot deletion can be conducted independently among VMs.
%sharing in our approach is limited to only a small dataset, we can greatly 
%on intra-VM sharing rather than inter-VM sharing. Then

\item Second,  we propose an {\em approximate} deletion strategy to trade deletion accuracy for speed and
resource usage. Since the VM size is highly skewed in practice,  a large VM may still require
a substantial amount of memory  to facilitate the mark-and-sweep process of data chunks used within all snapshots
of this VM. 
Our approximate deletion method sacrifices a small percent of storage leakage
to efficiently identify unused chunks using a per-snapshot bloom-filter.
\end{itemize}
The simplified deletion algorithm contains three aspects.
\begin{itemize}
\item {\bf Computation for snapshot reference summary.}
Every time there is a new snapshot created,
we compute a Bloom-filter with $z$ bits as the reference summary vector for all non-PDS chunks used 
in this snapshot.
The items we put into the summary vector are all the references appearing in the metadata of the snapshot.
For each VM we preset the vector size according to estimated VM image size;
given $h$ snapshots stored for a VM, there are $h$ summary vectors maintained.
We adjust the summary vector size and recompute the vectors if the VM size changes substantially over time.
This can be done during periodic leakage repair described below.
